---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

In this study, we employed a Generalized Linear Model (GLM) to predict the outcome of the upcoming US presidential election. Specifically, we used polling data to forecast the vote percentage for each candidate. The model was developed using the poll-of-polls methodology, which aggregates data across multiple polls, allowing for a more robust prediction by reducing individual poll biases. [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the vote percentage for candidate $i$. We modeled $y_i$ using a Gaussian family distribution, assuming a linear relationship between the response and the predictors. The predictors included the poll score ($pollscore$), the log-transformed sample size ($log_sample_size$), and state identifier ($state$).

\begin{align}
y_i|\mu_i, \sigma &\sim \text{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_1 \cdot \text{pollscore}_i + \beta_2 \cdot \text{log\_sample\_size}_i + \gamma_{\text{state}_i} \\
\alpha &\sim \text{Normal}(0, 2.5) \\
\beta_1, \beta_2 &\sim \text{Normal}(0, 2.5) \\
\gamma &\sim \text{Normal}(0, 2.5) \\
\sigma &\sim \text{Exponential}(1)
\end{align}

The model was implemented using the rstanarm package in R, utilizing default priors for all parameters.

To assess the model’s performance, we performed an extensive diagnostic evaluation. This included plotting residuals, analyzing their distribution, and calculating key goodness-of-fit metrics.

Residual vs. Fitted Plot: We examined the residuals versus fitted values to assess any systematic patterns that would indicate non-linearity or heteroscedasticity. The residuals appeared to be randomly distributed, suggesting that the GLM assumptions of linearity and homoscedasticity were largely met.

Normal Q-Q Plot: A Q-Q plot was used to evaluate the normality of the residuals. Most of the residuals fell along the reference line, indicating approximate normality, though some deviations at the tails suggested the presence of outliers.

Residual Histogram: The histogram of residuals showed a roughly bell-shaped distribution, further supporting the assumption of normality.

Cook's Distance: We computed Cook's distance to identify any influential observations. A small number of data points had higher Cook's distance values, suggesting they may have had a disproportionate influence on the model. These observations were further investigated, but no significant issues were found that warranted their removal.

R-squared and AIC: The R-squared value indicated a reasonable fit, explaining a significant portion of the variance in the vote percentage. The Akaike Information Criterion (AIC) was used to compare different models during stepwise feature selection, helping us to balance model complexity with predictive performance.

### Model justification



# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References
